---
title: "Forward - Age vs Performance"
format: pdf
---

## Filters of Analysis

We have limited our data set to the first 100 players who meet the criteria of playing at least 500 minutes during the season.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)

# Load the dataset
playerstats <- read.csv("nba-player-stats-2021.csv")

# Filter for Forwards with at least 500 minutes played
forward_data <- playerstats |> 
  filter(pos == 'PF' | pos == 'SF') |> 
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, orb, drb, ast, fgpercent, tov)

# Normalize the tracked statistics by minutes played
cols_to_normalize <- c("pts", "orb", "drb", "ast", "fgpercent", "tov")

for (colname in cols_to_normalize) {
  forward_data[[colname]] <- round(forward_data[[colname]] / forward_data[["mp"]], 6)
}

# Limit to the first 100 players
forward_data <- head(forward_data, 100)
```

## Performance Visualizations

Each point on the graphs below represents an individual player, categorized by their specific forward designation (PF or SF).

### 1. Points Scored vs Age

```{r}
ggplot(data = forward_data, mapping = aes(x = age, y = pts, color = pos)) +
  geom_point() +
  labs(
    x = "Age",
    y = "Points / Minute",
    title = "Age vs Points per Minute",
    color = "Position"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

*Points scored per minute relative to player age, showing scoring efficiency.*

### 2. Offensive Rebounds vs Age

```{r}
ggplot(data = forward_data, mapping = aes(x = age, y = orb, color = pos)) +
  geom_point() +
  labs(
    x = "Age",
    y = "Offensive Rebounds / Minute",
    title = "Age vs Offensive Rebounds per Minute",
    color = "Position"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

*Tracking how aggressive forwards are on the offensive glass as they age.*

### 3. Defensive Rebounds vs Age

```{r}
ggplot(data = forward_data, mapping = aes(x = age, y = drb, color = pos)) +
  geom_point() +
  labs(
    x = "Age",
    y = "Defensive Rebounds / Minute",
    title = "Age vs Defensive Rebounds per Minute",
    color = "Position"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

*Defensive rebounding efficiency across different age groups.*

### 4. Assists vs Age

```{r}
ggplot(data = forward_data, mapping = aes(x = age, y = ast, color = pos)) +
  geom_point() +
  labs(
    x = "Age",
    y = "Assists / Minute",
    title = "Age vs Assists per Minute",
    color = "Position"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

*Analyzing the playmaking ability of forwards relative to their experience level.*

### 5. Field Goal Percentage vs Age

```{r}
ggplot(data = forward_data, mapping = aes(x = age, y = fgpercent, color = pos)) +
  geom_point() +
  labs(
    x = "Age",
    y = "FG% / Minute",
    title = "Age vs Field Goal Percentage per Minute",
    color = "Position"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

*Shooting efficiency normalized by playtime.*

### 6. Turnovers vs Age

```{r}
ggplot(data = forward_data, mapping = aes(x = age, y = tov, color = pos)) +
  geom_point() +
  labs(
    x = "Age",
    y = "Turnovers / Minute",
    title = "Age vs Turnovers per Minute",
    color = "Position"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

*Ball security trends in forwards as they progress through their careers.*

## Table of all 100 Forwards
```{r}
forward_data |> 
  knitr::kable() |> 
  kableExtra::kable_classic()
```
