---
title: "Forward - Age vs Performance"
format: pdf
---

## Description of Analysis

In this analysis, we investigate how age impacts the performance of players in the **Forward** position (including both Power Forwards and Small Forwards). To ensure a fair comparison between starters and bench players, we normalize all statistics by dividing them by total minutes played (`mp`). This creates a ratio that represents skill and efficiency per minute, eliminating the bias that comes from simply having more time on the court.

We have limited our data set to the first 100 players who meet the criteria of playing at least 500 minutes during the season.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)

# Load the dataset
playerstats <- read.csv("nba-player-stats-2021.csv")

# Filter for Forwards with at least 500 minutes played
forward_data <- playerstats |> 
  filter(pos == 'PF' | pos == 'SF') |> 
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, orb, drb, ast, fgpercent, tov)

# Normalize the tracked statistics by minutes played
cols_to_normalize <- c("pts", "orb", "drb", "ast", "fgpercent", "tov")

for (colname in cols_to_normalize) {
  forward_data[[colname]] <- round(forward_data[[colname]] / forward_data[["mp"]], 6)
}

# Limit to the first 100 players
forward_data <- head(forward_data, 100)
```

## Performance Visualizations

Each point on the graphs below represents an individual player, categorized by their specific forward designation (PF or SF).

### 1. Points Scored vs Age

```{r}
ggplot(data = forward_data, mapping = aes(x = age, y = pts, color = pos)) +
  geom_point() +
  labs(
    x = "Age",
    y = "Points / Minute",
    title = "Age vs Points per Minute",
    color = "Position"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

*Points scored per minute relative to player age, showing scoring efficiency.*

### 2. Offensive Rebounds vs Age

```{r}
ggplot(data = forward_data, mapping = aes(x = age, y = orb, color = pos)) +
  geom_point() +
  labs(
    x = "Age",
    y = "Offensive Rebounds / Minute",
    title = "Age vs Offensive Rebounds per Minute",
    color = "Position"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

*Tracking how aggressive forwards are on the offensive glass as they age.*

### 3. Defensive Rebounds vs Age

```{r}
ggplot(data = forward_data, mapping = aes(x = age, y = drb, color = pos)) +
  geom_point() +
  labs(
    x = "Age",
    y = "Defensive Rebounds / Minute",
    title = "Age vs Defensive Rebounds per Minute",
    color = "Position"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

*Defensive rebounding efficiency across different age groups.*

### 4. Assists vs Age

```{r}
ggplot(data = forward_data, mapping = aes(x = age, y = ast, color = pos)) +
  geom_point() +
  labs(
    x = "Age",
    y = "Assists / Minute",
    title = "Age vs Assists per Minute",
    color = "Position"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

*Analyzing the playmaking ability of forwards relative to their experience level.*

### 5. Field Goal Percentage vs Age

```{r}
ggplot(data = forward_data, mapping = aes(x = age, y = fgpercent, color = pos)) +
  geom_point() +
  labs(
    x = "Age",
    y = "FG% / Minute",
    title = "Age vs Field Goal Percentage per Minute",
    color = "Position"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

*Shooting efficiency normalized by playtime.*

### 6. Turnovers vs Age

```{r}
ggplot(data = forward_data, mapping = aes(x = age, y = tov, color = pos)) +
  geom_point() +
  labs(
    x = "Age",
    y = "Turnovers / Minute",
    title = "Age vs Turnovers per Minute",
    color = "Position"
  ) + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

*Ball security trends in forwards as they progress through their careers.*

## Discussion of Findings

Upon analyzing the data for Forwards, several key trends emerge:

**Efficiency and Age:** Scoring efficiency (Points/Minute) remains relatively stable throughout the mid-20s, with a slight peak in the late 20s. This suggests that Forwards maintain their scoring utility even as their pure athleticism might begin to transition.

**Rebounding Trends:** We notice that Power Forwards (PF) generally maintain higher offensive and defensive rebounding ratios compared to Small Forwards (SF) across almost all ages, which aligns with their traditional roles closer to the basket.

**Playmaking Growth:** There appears to be a slight upward trend in assists per minute as players reach their early 30s. This may indicate that veteran forwards develop better court vision and passing skills to compensate for changes in physical speed.

**Ball Security:** Turnovers per minute show a wide variance among younger players, while older players tend to cluster more tightly, potentially indicating more disciplined play or more specialized roles as they age.

Overall, the normalization by minutes played allows us to see that a Forward's value is not just a product of how long they are on the floor, but how effectively they use that time at different stages of their career.
