---
title: "NBA Age vs. Performance: A Cross-Position Analysis"
author: "Rishvanth Amsaraj, Vishnu Bommisetty, Samarth Puri"
date: "2026-04-28"
format:
  pdf:
    toc: true
    number-sections: true
    fig-pos: "H"
    geometry: margin=1in
    fontsize: 11pt
execute:
  echo: false
  warning: false
  message: false
---

\newpage

```{r setup}
# Style Guide: Tidyverse Style Guide (https://style.tidyverse.org/)
# Primary Authors: Rishvanth Amsaraj, Vishnu Bommisetty, Samarth Puri
# Reviewer: All team members

library(tidyverse)
library(ggplot2)
library(patchwork)
library(knitr)
library(kableExtra)

# Helper: builds a single scatter plot panel
make_panel <- function(df, y_var, y_label, panel_title) {
  ggplot(df, aes(x = age, y = .data[[y_var]], color = pos, shape = pos)) +
    geom_point(alpha = 0.75, size = 1.6) +
    labs(
      title  = panel_title,
      x      = "Age",
      y      = y_label,
      color  = "Position",
      shape  = "Position"
    ) +
    theme_minimal(base_size = 9) +
    theme(
      plot.title      = element_text(hjust = 0.5, size = 9, face = "bold"),
      legend.position = "bottom",
      legend.key.size = unit(0.35, "cm"),
      legend.text     = element_text(size = 7),
      legend.title    = element_text(size = 7)
    )
}

# Helper: assembles a 2x3 patchwork figure for a position group
make_position_figure <- function(df) {
  p1 <- make_panel(df, "pts",       "Points / Min",        "Points")
  p2 <- make_panel(df, "ast",       "Assists / Min",       "Assists")
  p3 <- make_panel(df, "orb",       "Off. Rebounds / Min", "Off. Rebounds")
  p4 <- make_panel(df, "drb",       "Def. Rebounds / Min", "Def. Rebounds")
  p5 <- make_panel(df, "fgpercent", "FG% / Min",           "Field Goal %")
  p6 <- make_panel(df, "tov",       "Turnovers / Min",     "Turnovers")

  ((p1 | p2) / (p3 | p4) / (p5 | p6)) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
}
```

# Introduction

This report looks at how NBA player age relates to on-court performance across three position groups:Guards (Point Guards and Shooting Guards), Forwards (Power Forwards and Small Forwards),and Centers. We use data from the 2020–21 NBA season (Basketball Reference, 2021) and track six statistics: points, assists, offensive rebounds, defensive rebounds, field goal percentage, and turnovers.

Rather than using raw totals, we divide every stat by minutes played. This gives a per-minute rate that makes it fair to compare players who get very different amounts of playing time. The goal is to explore how these per-minute rates change with age for each position group.

# Data and Methods

## Data Source

The data comes from `nba-player-stats-2021.csv`, a publicly available dataset
of NBA player statistics for the 2020--21 season (Basketball Reference, 2021).
Each case is a single NBA player's season record. The attributes used are:
`age` (years), `pos` (position: PG, SG, PF, SF, or C), `mp` (total minutes
played), `pts` (points), `ast` (assists), `orb` (offensive rebounds), `drb`
(defensive rebounds), `fgpercent` (field goal percentage, 0 to 1), and `tov`
(turnovers). After normalization, all counting stats are rates per minute.

## Player Selection Criteria

Players with fewer than 500 minutes played were excluded to avoid unreliable
per-minute rates. Each position group is capped at the first 100 qualifying
players.

## FAIR and CARE Principles

The dataset satisfies the **FAIR** principles: it is publicly Findable and
Accessible via Basketball Reference, stored in a standard CSV format
(Interoperable), with clear column naming (Reusable). Applying the **CARE**
principles is more challenging, as these guidelines target Indigenous data
sovereignty. One challenge is that players had no input into how their
statistics are collected or used in third-party analyses.

## Per-Minute Normalization

Each counting stat (points, assists, rebounds, turnovers) and field goal
percentage are divided by minutes played and rounded to six decimal places.
The same process is applied consistently across all three position groups.

## Analysis Paradigm {#sec-paradigm}

This analysis uses an **Exploratory Data Analysis (EDA)** approach (Tukey,
1977) --- we use visualizations to look for patterns rather than test a
specific hypothesis. All code is written using the **Tidyverse** (Wickham
et al., 2019), chosen because the pipe operator and `ggplot2` produce
readable, step-by-step code.

## PCIP Plan Summary

The full PCIP plan is in Appendix 1. The four phases were: **Plan** (define
the question and steps), **Code** (write the pipeline following the Tidyverse
Style Guide), **Implement** (run and verify output), and **Present** (write
the report with figures and appendices).

# Guards: Point Guards and Shooting Guards

```{r guards-data}
# Primary Author: Samarth Puri | Reviewer: Rishvanth Amsaraj

playerstats <- read.csv("nba-player-stats-2021.csv")

# Check structure and dimensions immediately after import
glimpse(playerstats)

# Filter to PG and SG with > 500 minutes; select analysis columns
guard_data <- playerstats |>
  filter(pos == "PG" | pos == "SG") |>
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, ast, orb, drb, fgpercent, tov)

# Normalize each statistic by minutes played
cols_to_normalize <- c("pts", "ast", "orb", "drb", "fgpercent", "tov")
guard_data <- guard_data |>
  mutate(across(all_of(cols_to_normalize), ~ round(. / mp, 6)))

guard_data <- head(guard_data, 100)
```

Point Guards (PG) are mainly responsible for running the offense and setting up
plays, while Shooting Guards (SG) focus more on scoring from the perimeter.
Both spend most of their time on the outside of the court, so we would expect
assists to be higher for PGs and scoring to be a priority for SGs.
Figure\ @fig-guards shows how each per-minute stat looks across different ages
for both guard types.

```{r plot-guards}
#| label: fig-guards
#| fig-cap: "Age versus six per-minute performance metrics for NBA Guards (PG and SG) with at least 500 minutes played in the 2020--21 season. Each point represents one player, colored by position (PG = Point Guard, SG = Shooting Guard)."
#| fig-alt: "A 2-by-3 grid of scatter plots. The x-axis of every panel shows player age roughly 18 to 38. The y-axes show from left to right and top to bottom: points per minute, assists per minute, offensive rebounds per minute, defensive rebounds per minute, field goal percentage per minute, and turnovers per minute. Points are colored by guard subtype."
#| fig-width: 7
#| fig-height: 8.5

make_position_figure(guard_data)
```

Looking at Figure\ @fig-guards, midway through their twenties into their thirties, guards display varied scoring rates per minute, seen clearly in both point and shooting categories. While one might expect sharp differences, the data shows overlap instead. Notably, assists diverge - point guards consistently record higher numbers, matching expectations of them as primary creators. Unlike forwards, these backcourt players rarely collect offensive boards, given their positioning farther from the rim. Among younger cohorts, turnover counts scatter widely, yet they tighten with advancing age, perhaps due to greater experience shaping consistent responsibilities. Efficiency, measured by field goal accuracy, stays mostly flat across years for each group, without distinct peaks or declines. Table\ @tbl-guard-summary shows mean and standard deviation for each per-minute stat across all 100 guards, and Table\ @tbl-guards lists the full player-level data.

## Summary Statistics for Guards

```{r tbl-guard-summary}
#| label: tbl-guard-summary
#| tbl-cap: "Mean and standard deviation of per-minute statistics for the 100 guards included in the analysis."
guard_data |>
  select(age, all_of(cols_to_normalize)) |>
  summarise(across(everything(), list(Mean = mean, SD = sd))) |>
  pivot_longer(
    everything(),
    names_to  = c("Variable", ".value"),
    names_sep = "_"
  ) |>
  kable(booktabs = TRUE, digits = 4, linesep = "") |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 8
  )
```

## Table of Guards Used in Analysis

```{r tbl-guards}
#| label: tbl-guards
#| tbl-cap: "Per-minute statistics for all 100 guards (PG and SG) included in the analysis."
guard_data |>
  kable(booktabs = TRUE) |>
  kable_styling(
    latex_options = c("scale_down", "hold_position"),
    font_size = 7
  )
```

# Forwards: Power Forwards and Small Forwards

```{r forwards-data}
# Primary Author: Rishvanth Amsaraj | Reviewer: Vishnu Bommisetty

# Filter to PF and SF with > 500 minutes; select analysis columns
forward_data <- playerstats |>
  filter(pos == "PF" | pos == "SF") |>
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, ast, orb, drb, fgpercent, tov)

# Normalize each statistic by minutes played
forward_data <- forward_data |>
  mutate(across(all_of(cols_to_normalize), ~ round(. / mp, 6)))

forward_data <- head(forward_data, 100)
```

Power Forwards (PF) tend to play closer to the basket and focus on rebounding
and inside scoring. Small Forwards (SF) are more versatile --- they can score
from the perimeter and also get to the rim. Because these two roles use the
court differently, it is interesting to see how age affects each one.
Figure\ @fig-forwards shows the per-minute stats for both forward types.

```{r plot-forwards}
#| label: fig-forwards
#| fig-cap: "Age versus six per-minute performance metrics for NBA Forwards (PF and SF) with at least 500 minutes played in the 2020--21 season. Each point represents one player, colored by position (PF = Power Forward, SF = Small Forward)."
#| fig-alt: "A 2-by-3 grid of scatter plots. The x-axis of every panel shows player age roughly 18 to 38. The y-axes show from left to right and top to bottom: points per minute, assists per minute, offensive rebounds per minute, defensive rebounds per minute, field goal percentage per minute, and turnovers per minute. Points are colored by forward subtype."
#| fig-width: 7
#| fig-height: 8.5

make_position_figure(forward_data)
```

Figure\ @fig-forwards shows that Scoring each minute holds steady through the mid-twenties, though a bit more often peaks between ages twenty-six and thirty. While assists generally run lower for forwards compared to guards, several small forwards past age twenty-five post notably higher figures. What stands apart is rebounding - power forwards grab boards far more frequently than small forwards, regardless of age, without narrowing much over time. That pattern hints role shapes rebound output stronger than aging does. Younger players show wider turnover swings early on; these tend to even out later, echoing trends seen earlier in guard data. Table\ @tbl-forward-summary summarizes the per-minute statistics and Table\ @tbl-forwards lists all 100 players.

## Summary Statistics for Forwards

```{r tbl-forward-summary}
#| label: tbl-forward-summary
#| tbl-cap: "Mean and standard deviation of per-minute statistics for the 100 forwards included in the analysis."
forward_data |>
  select(age, all_of(cols_to_normalize)) |>
  summarise(across(everything(), list(Mean = mean, SD = sd))) |>
  pivot_longer(
    everything(),
    names_to  = c("Variable", ".value"),
    names_sep = "_"
  ) |>
  kable(booktabs = TRUE, digits = 4, linesep = "") |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 8
  )
```

## Table of Forwards Used in Analysis

```{r tbl-forwards}
#| label: tbl-forwards
#| tbl-cap: "Per-minute statistics for all 100 forwards (PF and SF) included in the analysis."
forward_data |>
  kable(booktabs = TRUE) |>
  kable_styling(
    latex_options = c("scale_down", "hold_position"),
    font_size = 7
  )
```

# Centers

```{r centers-data}
# Primary Author: Vishnu Bommisetty | Reviewer: Rishvanth Amsaraj

# Filter to Centers with > 500 minutes; select analysis columns
center_data <- playerstats |>
  filter(pos == "C") |>
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, ast, orb, drb, fgpercent, tov)

# Normalize each statistic by minutes played
center_data <- center_data |>
  mutate(across(all_of(cols_to_normalize), ~ round(. / mp, 6)))

center_data <- head(center_data, 100)
```

Centers generally play near the basket. Their main jobs are rebounding,
protecting the rim, and scoring inside. Because there is only one position
label (C), all points in Figure\ @fig-centers are the same color, so the focus
is entirely on how age relates to each stat.

```{r plot-centers}
#| label: fig-centers
#| fig-cap: "Age versus six per-minute performance metrics for NBA Centers (C) with at least 500 minutes played in the 2020--21 season. Each point represents one player. All players share the same position label (C)."
#| fig-alt: "A 2-by-3 grid of scatter plots. The x-axis of every panel shows player age roughly 18 to 38. The y-axes show from left to right and top to bottom: points per minute, assists per minute, offensive rebounds per minute, defensive rebounds per minute, field goal percentage per minute, and turnovers per minute. All points are the same color since all players are Centers."
#| fig-width: 7
#| fig-height: 8.5

make_position_figure(center_data)
```

Looking at Figure @fig-centers, scoring rates for centers stay close together - most fall from 0.4 to 0.8 points each minute - with little shift across ages. Because creating shots isn’t typically their role, it follows that assist numbers sit below those of other positions. Unlike guards or forwards, these players pull down more rebounds, as seen when comparing with Figures @fig-guards and @fig-forwards; yet a small decline appears in defensive boards past the early years.

By the third decade of life, mobility tends to decrease slightly. Although turnover rates begin elevated among new center players, they tend to decline with experience. This trend mirrors what is observed in back court and wing positions alike. Table\ @tbl-center-summary provides summary statistics and Table\ @tbl-centers lists all centers used.

## Summary Statistics for Centers

```{r tbl-center-summary}
#| label: tbl-center-summary
#| tbl-cap: "Mean and standard deviation of per-minute statistics for the centers included in the analysis."
center_data |>
  select(age, all_of(cols_to_normalize)) |>
  summarise(across(everything(), list(Mean = mean, SD = sd))) |>
  pivot_longer(
    everything(),
    names_to  = c("Variable", ".value"),
    names_sep = "_"
  ) |>
  kable(booktabs = TRUE, digits = 4, linesep = "") |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 8
  )
```

## Table of Centers Used in Analysis

```{r tbl-centers}
#| label: tbl-centers
#| tbl-cap: "Per-minute statistics for all centers (C) included in the analysis."
center_data |>
  kable(booktabs = TRUE) |>
  kable_styling(
    latex_options = c("scale_down", "hold_position"),
    font_size = 7
  )
```

# Cross-Position Discussion

Surprisingly, among these roles, where you play shapes performance more than how old you are. Rebounding? That belongs to centers first, then forwards, leaving guards behind - no matter the years played. Assist numbers jump highest with point guards, pulling well ahead of others across positions. Age changes little; those gaps stay fixed over time.

With advancing years, dispersion patterns shift noticeably across positions. While younger athletes display broader performance ranges - some excelling quickly, others lagging - the data for older ones pulls closer together. This tightening likely emerges as experience shapes consistent contributions over time.

Surprisingly steady scoring rates emerge per minute across age groups when focusing on those with 500 minutes played. That pattern probably appears since the minimum playing time cuts out irregular participants - often younger prospects still growing or aging ones near the edge of squads.

One season’s worth of data limits how far conclusions can stretch. Top performers were pulled from a non-random selection, which may tilt results. Player roles within team setups, along with absences due to injury, influence output per minute played. Patterns seen here might hold up under stricter analysis later. Testing them again would help confirm what shows now.

# Reflection

Surprisingly, diving into this work revealed how crucial stat normalization really is. Without context, raw counts often distort reality - someone logging heavier playing time naturally racks up bigger figures; scaling them down per minute brings things into clearer view.

Putting together three distinct files into a single document raised questions about uniformity. Since each file originally applied its own naming and layout choices, aligning them became essential for meaningful analysis. To manage this, common tools - like `make_panel` and `make_position_figure` - were introduced. Rather than repeat plotting logic across scripts, these functions allowed repetition without redundancy. Consistency improved simply by centralizing what was once scattered.

Putting the code into the appendix helped keep the main section clear. What resulted was a report that flowed better, thanks to separation of details. Readers found it easier to follow when results stayed up front. The structure gave space for analysis without distraction. Length stayed the same, yet clarity improved simply by moving parts around.

\newpage

# Author Contributions {.unnumbered}

**Rishvanth Amsaraj** --- Forwards analysis (data filtering, normalization, and
figure), combining all three position files into the final QMD, patchwork
figure layout, shared helper functions, narrative text, and PCIP documentation.

**Vishnu Bommisetty** --- Centers analysis (data filtering, normalization, and
figure), and the repository README.md file.

**Samarth Puri** --- Guards analysis (data filtering, normalization, and
figure), and the repository PLAN.md file.

The overall analysis approach, filtering criteria, normalization method, and
report structure were decided together as a team.

\newpage

# Appendix 1: PCIP Written Plan {.unnumbered}

The PCIP System (Plan, Code, Implement, Present) was applied to both the data
analysis and the coding work in this project.

## Phase 1 --- Plan {.unnumbered}

**Research question:** How does player age relate to six per-minute performance
statistics across three NBA position groups (Guards, Forwards, Centers)?

**Data source:** `nba-player-stats-2021.csv` --- 2020--21 NBA season, one row
per player.

**Issues identified:**

- Stats are season totals, so players with more minutes look better without
  normalization. Dividing by `mp` fixes this.
- Some players have very few minutes, making their per-minute rates unreliable.
  A 500-minute filter removes these cases.
- The three original files used different variable names and inconsistent
  formatting, which needed to be unified.

**Steps planned:**

1. Load `nba-player-stats-2021.csv` once in the setup chunk.
2. Write shared `make_panel()` and `make_position_figure()` helpers to avoid
   repeating plot code for each position group.
3. For each group: filter by position and `mp > 500`, select the needed
   columns, divide stats by `mp`, and cap at 100 rows.
4. Produce a 2x3 patchwork figure per group with a shared legend.
5. Write narrative for each section and a cross-position discussion.
6. Set up YAML to match the assignment template (TOC, numbered sections,
   1-inch margins, 11pt font, `echo: false` globally).
7. Move all code to a Code Appendix.

## Phase 2 --- Code {.unnumbered}

All code follows the **Tidyverse Style Guide** (https://style.tidyverse.org/).
Key practices:

- Snake\_case for all variable names.
- Pipe operator (`|>`) for data transformation steps.
- Spaces around operators and after commas.
- One `ggplot2` verb per line with consistent indentation.
- Each chunk starts with a comment naming the primary author and reviewer.

## Phase 3 --- Implement {.unnumbered}

The code was run in R through Quarto. We checked that filtering produced the
right rows for each position group, that all six figures rendered correctly,
and that cross-references resolved properly in the PDF.

## Phase 4 --- Present {.unnumbered}

The final PDF was produced by rendering `NBA_Age_vs_Performance.qmd`. The
report includes a table of contents, three cross-referenced figures with
captions and alt text, narrative in every section, a cross-position discussion,
a reflection, author contributions, this PCIP plan, and a code appendix.

\newpage

# Appendix 2: Code Appendix {.unnumbered}

```{r code-appendix, echo=TRUE, eval=FALSE}
# ==========================================================
# Style Guide: Tidyverse Style Guide
# https://style.tidyverse.org/
# ==========================================================

# ── Libraries ──────────────────────────────────────────────
library(tidyverse)
library(ggplot2)
library(patchwork)
library(knitr)
library(kableExtra)

# ── Helper: single scatter panel ──────────────────────────
# Primary Authors: Rishvanth Amsaraj, Vishnu Bommisetty, Samarth Puri
# Reviewer: All team members
make_panel <- function(df, y_var, y_label, panel_title) {
  ggplot(df, aes(x = age, y = .data[[y_var]], color = pos, shape = pos)) +
    geom_point(alpha = 0.75, size = 1.6) +
    labs(
      title  = panel_title,
      x      = "Age",
      y      = y_label,
      color  = "Position",
      shape  = "Position"
    ) +
    theme_minimal(base_size = 9) +
    theme(
      plot.title      = element_text(hjust = 0.5, size = 9, face = "bold"),
      legend.position = "bottom",
      legend.key.size = unit(0.35, "cm"),
      legend.text     = element_text(size = 7),
      legend.title    = element_text(size = 7)
    )
}

# ── Helper: 2x3 patchwork figure for a position group ─────
# Primary Authors: Rishvanth Amsaraj, Vishnu Bommisetty, Samarth Puri
# Reviewer: All team members
make_position_figure <- function(df) {
  p1 <- make_panel(df, "pts",       "Points / Min",        "Points")
  p2 <- make_panel(df, "ast",       "Assists / Min",       "Assists")
  p3 <- make_panel(df, "orb",       "Off. Rebounds / Min", "Off. Rebounds")
  p4 <- make_panel(df, "drb",       "Def. Rebounds / Min", "Def. Rebounds")
  p5 <- make_panel(df, "fgpercent", "FG% / Min",           "Field Goal %")
  p6 <- make_panel(df, "tov",       "Turnovers / Min",     "Turnovers")

  ((p1 | p2) / (p3 | p4) / (p5 | p6)) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
}

# ==========================================================
# Section 3: Guards (PG and SG)
# Primary Author: Samarth Puri | Reviewer: Rishvanth Amsaraj
# ==========================================================
playerstats <- read.csv("nba-player-stats-2021.csv")

# Check structure and dimensions immediately after import
glimpse(playerstats)

# Filter to guards with > 500 mp; select analysis columns
guard_data <- playerstats |>
  filter(pos == "PG" | pos == "SG") |>
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, ast, orb, drb, fgpercent, tov)

# Normalize each statistic by minutes played using across()
cols_to_normalize <- c("pts", "ast", "orb", "drb", "fgpercent", "tov")
guard_data <- guard_data |>
  mutate(across(all_of(cols_to_normalize), ~ round(. / mp, 6)))
guard_data <- head(guard_data, 100)

make_position_figure(guard_data)

# Summary statistics (mean and SD per variable)
guard_data |>
  select(age, all_of(cols_to_normalize)) |>
  summarise(across(everything(), list(Mean = mean, SD = sd))) |>
  pivot_longer(
    everything(),
    names_to  = c("Variable", ".value"),
    names_sep = "_"
  ) |>
  kable(booktabs = TRUE, digits = 4, linesep = "") |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 8
  )

# Full player-level data table
guard_data |>
  kable(booktabs = TRUE) |>
  kable_styling(
    latex_options = c("scale_down", "hold_position"),
    font_size = 7
  )

# ==========================================================
# Section 4: Forwards (PF and SF)
# Primary Author: Rishvanth Amsaraj | Reviewer: Vishnu Bommisetty
# ==========================================================

# Filter to forwards with > 500 mp; select analysis columns
forward_data <- playerstats |>
  filter(pos == "PF" | pos == "SF") |>
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, ast, orb, drb, fgpercent, tov)

# Normalize each statistic by minutes played using across()
forward_data <- forward_data |>
  mutate(across(all_of(cols_to_normalize), ~ round(. / mp, 6)))
forward_data <- head(forward_data, 100)

make_position_figure(forward_data)

# Summary statistics (mean and SD per variable)
forward_data |>
  select(age, all_of(cols_to_normalize)) |>
  summarise(across(everything(), list(Mean = mean, SD = sd))) |>
  pivot_longer(
    everything(),
    names_to  = c("Variable", ".value"),
    names_sep = "_"
  ) |>
  kable(booktabs = TRUE, digits = 4, linesep = "") |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 8
  )

# Full player-level data table
forward_data |>
  kable(booktabs = TRUE) |>
  kable_styling(
    latex_options = c("scale_down", "hold_position"),
    font_size = 7
  )

# ==========================================================
# Section 5: Centers (C)
# Primary Author: Vishnu Bommisetty | Reviewer: Rishvanth Amsaraj
# ==========================================================

# Filter to centers with > 500 mp; select analysis columns
center_data <- playerstats |>
  filter(pos == "C") |>
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, ast, orb, drb, fgpercent, tov)

# Normalize each statistic by minutes played using across()
center_data <- center_data |>
  mutate(across(all_of(cols_to_normalize), ~ round(. / mp, 6)))
center_data <- head(center_data, 100)

make_position_figure(center_data)

# Summary statistics (mean and SD per variable)
center_data |>
  select(age, all_of(cols_to_normalize)) |>
  summarise(across(everything(), list(Mean = mean, SD = sd))) |>
  pivot_longer(
    everything(),
    names_to  = c("Variable", ".value"),
    names_sep = "_"
  ) |>
  kable(booktabs = TRUE, digits = 4, linesep = "") |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 8
  )

# Full player-level data table
center_data |>
  kable(booktabs = TRUE) |>
  kable_styling(
    latex_options = c("scale_down", "hold_position"),
    font_size = 7
  )
```

## References {.unnumbered}

Basketball Reference. (2021). *NBA player stats --- 2020--21 season* [Dataset].
Retrieved from https://www.basketball-reference.com

Tukey, J. W. (1977). *Exploratory data analysis*. Addison-Wesley.

Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D., François, R.,
Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T. L.,
Miller, E., Bache, S. M., Müller, K., Ooms, J., Robinson, D., Seidel, D. P.,
Spinu, V., Takahashi, K., Vaughan, D., Wilke, C., Woo, K., & Yutani, H. (2019).
Welcome to the tidyverse. *Journal of Open Source Software*, *4*(43), 1686.
https://doi.org/10.21105/joss.01686
