---
title: "NBA Age vs. Performance: A Cross-Position Analysis"
author: "Rishvanth Amsaraj, Vishnu Bombesetti, Samarth Puri"
date: "2026-04-28"
format:
  pdf:
    toc: true
    number-sections: true
    fig-pos: "H"
    geometry: margin=1in
    fontsize: 11pt
execute:
  echo: false
  warning: false
  message: false
---

\newpage

```{r setup}
# Style Guide: Tidyverse Style Guide (https://style.tidyverse.org/)
# Primary Authors: Rishvanth Amsaraj, Vishnu Bombesetti, Samarth Puri
# Reviewer: All team members

library(tidyverse)
library(ggplot2)
library(patchwork)
library(knitr)
library(kableExtra)

# Helper: builds a single scatter plot panel
make_panel <- function(df, y_var, y_label, panel_title) {
  ggplot(df, aes(x = age, y = .data[[y_var]], color = pos)) +
    geom_point(alpha = 0.75, size = 1.6) +
    labs(
      title  = panel_title,
      x      = "Age",
      y      = y_label,
      color  = "Position"
    ) +
    theme_minimal(base_size = 9) +
    theme(
      plot.title      = element_text(hjust = 0.5, size = 9, face = "bold"),
      legend.position = "bottom",
      legend.key.size = unit(0.35, "cm"),
      legend.text     = element_text(size = 7),
      legend.title    = element_text(size = 7)
    )
}

# Helper: assembles a 2x3 patchwork figure for a position group
make_position_figure <- function(df) {
  p1 <- make_panel(df, "pts",       "Points / Min",        "Points")
  p2 <- make_panel(df, "ast",       "Assists / Min",       "Assists")
  p3 <- make_panel(df, "orb",       "Off. Rebounds / Min", "Off. Rebounds")
  p4 <- make_panel(df, "drb",       "Def. Rebounds / Min", "Def. Rebounds")
  p5 <- make_panel(df, "fgpercent", "FG% / Min",           "Field Goal %")
  p6 <- make_panel(df, "tov",       "Turnovers / Min",     "Turnovers")

  ((p1 | p2) / (p3 | p4) / (p5 | p6)) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
}
```

# Introduction

This report looks at how NBA player age relates to on-court performance across
three position groups: Guards (Point Guards and Shooting Guards), Forwards
(Power Forwards and Small Forwards), and Centers. We use data from the 2020--21
NBA season (Basketball Reference, 2021) and track six statistics: points,
assists, offensive rebounds, defensive rebounds, field goal percentage, and
turnovers.

Rather than using raw totals, we divide every stat by minutes played. This
gives a per-minute rate that makes it fair to compare players who get very
different amounts of playing time. The goal is to explore how these per-minute
rates change with age for each position group.

# Data and Methods

## Data Source

The data comes from `nba-player-stats-2021.csv`, a publicly available dataset
of NBA player statistics for the 2020--21 season (Basketball Reference, 2021).
It contains season totals for each player, and we use the following columns:
`age`, `pos`, `mp` (minutes played), `pts`, `ast`, `orb`, `drb`, `fgpercent`,
and `tov`.

## Player Selection Criteria

We only included players with more than 500 minutes played in the season. This
cuts out players with very limited playtime, whose per-minute numbers can be
all over the place. Within each position group the data is also capped at the
first 100 qualifying players to keep things consistent across groups.

## Per-Minute Normalization

Each counting stat (points, assists, rebounds, and turnovers) is divided by
minutes played and rounded to six decimal places. Field goal percentage is
handled the same way. The same process is applied to all three position groups
using the same code, so the comparison across groups is consistent.

## Analysis Paradigm {#sec-paradigm}

This analysis uses an **Exploratory Data Analysis (EDA)** approach (Tukey,
1977). We are not testing a specific hypothesis --- we are using visualizations
to look for patterns that might be worth studying more formally later. On the
coding side, everything is written using the **Tidyverse** (Wickham et al.,
2019), which we chose because the pipe operator and `ggplot2` make the code
easy to read and the steps easy to follow.

## PCIP Plan Summary

The full PCIP plan is in Appendix 1. The four phases were: **Plan** (define the
question and steps), **Code** (write the pipeline using the Tidyverse Style
Guide), **Implement** (run and check the output), and **Present** (write the
report with narrative, figures, and appendices).

# Guards: Point Guards and Shooting Guards

```{r guards-data}
# Primary Author: Samarth Puri | Reviewer: Rishvanth Amsaraj

playerstats <- read.csv("nba-player-stats-2021.csv")

# Filter to PG and SG with > 500 minutes; select analysis columns
guard_data <- playerstats |>
  filter(pos == "PG" | pos == "SG") |>
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, ast, orb, drb, fgpercent, tov)

# Normalize each statistic by minutes played
cols_to_normalize <- c("pts", "ast", "orb", "drb", "fgpercent", "tov")
for (colname in cols_to_normalize) {
  guard_data[[colname]] <- round(
    guard_data[[colname]] / guard_data[["mp"]], 6
  )
}

guard_data <- head(guard_data, 100)
```

Point Guards (PG) are mainly responsible for running the offense and setting up
plays, while Shooting Guards (SG) focus more on scoring from the perimeter.
Both spend most of their time on the outside of the court, so we would expect
assists to be higher for PGs and scoring to be a priority for SGs.
Figure\ @fig-guards shows how each per-minute stat looks across different ages
for both guard types.

```{r plot-guards}
#| label: fig-guards
#| fig-cap: "Age versus six per-minute performance metrics for NBA Guards (PG and SG) with at least 500 minutes played in the 2020--21 season. Each point represents one player, colored by position (PG = Point Guard, SG = Shooting Guard)."
#| fig-alt: "A 2-by-3 grid of scatter plots. The x-axis of every panel shows player age roughly 18 to 38. The y-axes show from left to right and top to bottom: points per minute, assists per minute, offensive rebounds per minute, defensive rebounds per minute, field goal percentage per minute, and turnovers per minute. Points are colored by guard subtype."
#| fig-width: 7
#| fig-height: 8.5

make_position_figure(guard_data)
```

Looking at Figure\ @fig-guards, scoring per minute is fairly spread out across
ages in the mid-20s to early 30s, with PGs and SGs both showing up across the
range. Assists tell a clearer story --- PGs are noticeably higher than SGs at
most ages, which lines up with their role as playmakers. Offensive rebounds are
low for both, since guards do not spend much time near the basket. Turnovers are
more spread out among younger players and seem to settle down as players get
older, which probably reflects that veterans are in more defined roles. Field
goal percentage does not show a clear trend with age for either guard type.

# Forwards: Power Forwards and Small Forwards

```{r forwards-data}
# Primary Author: Rishvanth Amsaraj | Reviewer: Vishnu Bombesetti

# Filter to PF and SF with > 500 minutes; select analysis columns
forward_data <- playerstats |>
  filter(pos == "PF" | pos == "SF") |>
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, ast, orb, drb, fgpercent, tov)

# Normalize each statistic by minutes played
for (colname in cols_to_normalize) {
  forward_data[[colname]] <- round(
    forward_data[[colname]] / forward_data[["mp"]], 6
  )
}

forward_data <- head(forward_data, 100)
```

Power Forwards (PF) tend to play closer to the basket and focus on rebounding
and inside scoring. Small Forwards (SF) are more versatile --- they can score
from the perimeter and also get to the rim. Because these two roles use the
court differently, it is interesting to see how age affects each one.
Figure\ @fig-forwards shows the per-minute stats for both forward types.

```{r plot-forwards}
#| label: fig-forwards
#| fig-cap: "Age versus six per-minute performance metrics for NBA Forwards (PF and SF) with at least 500 minutes played in the 2020--21 season. Each point represents one player, colored by position (PF = Power Forward, SF = Small Forward)."
#| fig-alt: "A 2-by-3 grid of scatter plots. The x-axis of every panel shows player age roughly 18 to 38. The y-axes show from left to right and top to bottom: points per minute, assists per minute, offensive rebounds per minute, defensive rebounds per minute, field goal percentage per minute, and turnovers per minute. Points are colored by forward subtype."
#| fig-width: 7
#| fig-height: 8.5

make_position_figure(forward_data)
```

Figure\ @fig-forwards shows that scoring per minute is fairly steady across the
mid-20s, with slightly more high scorers in the 26--30 age range. Assists are
lower for forwards than guards overall, but a few SFs in their late 20s stand
out with higher assist numbers. Rebounding is where PFs and SFs differ the
most: Power Forwards rebound at a higher rate than Small Forwards at almost
every age, and that gap does not close as players get older. This suggests
position role matters more than age when it comes to rebounding for forwards.
Turnovers are more variable among younger players and get more consistent with
age, similar to what we saw with guards.

# Centers

```{r centers-data}
# Primary Author: Vishnu Bombesetti | Reviewer: Rishvanth Amsaraj

# Filter to Centers with > 500 minutes; select analysis columns
center_data <- playerstats |>
  filter(pos == "C") |>
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, ast, orb, drb, fgpercent, tov)

# Normalize each statistic by minutes played
for (colname in cols_to_normalize) {
  center_data[[colname]] <- round(
    center_data[[colname]] / center_data[["mp"]], 6
  )
}

center_data <- head(center_data, 100)
```

Centers generally play near the basket. Their main jobs are rebounding,
protecting the rim, and scoring inside. Because there is only one position
label (C), all points in Figure\ @fig-centers are the same color, so the focus
is entirely on how age relates to each stat.

```{r plot-centers}
#| label: fig-centers
#| fig-cap: "Age versus six per-minute performance metrics for NBA Centers (C) with at least 500 minutes played in the 2020--21 season. Each point represents one player. All players share the same position label (C)."
#| fig-alt: "A 2-by-3 grid of scatter plots. The x-axis of every panel shows player age roughly 18 to 38. The y-axes show from left to right and top to bottom: points per minute, assists per minute, offensive rebounds per minute, defensive rebounds per minute, field goal percentage per minute, and turnovers per minute. All points are the same color since all players are Centers."
#| fig-width: 7
#| fig-height: 8.5

make_position_figure(center_data)
```

Figure\ @fig-centers shows that scoring per minute for centers is fairly
compact, with most players between 0.4 and 0.8 points per minute and no strong
age trend. Assists are the lowest of any position group, which makes sense
since setting up plays is not usually a center's main job. Rebounding is
noticeably higher than for guards or forwards (compare Figures\ @fig-guards and
@fig-forwards), and there is a slight drop in defensive rebounds after the early
30s, which could be related to slower movement as players age. Turnovers are a
bit higher for younger centers and settle down over time, consistent with the
same pattern seen for guards and forwards.

# Cross-Position Discussion

Looking across all three position groups, one pattern stands out: your position
matters more than your age when it comes to most per-minute stats. Centers
rebound more than forwards, who rebound more than guards, at every age. Guards
--- especially Point Guards --- have far more assists than any other group.
These differences hold steady as players get older.

What does seem to change with age is how spread out the numbers are. Younger
players at every position show a wider range of values --- some do really well
early, others struggle. As players get older the numbers tend to cluster more
tightly, which probably reflects that veterans settle into defined roles.

Scoring per minute is also surprisingly stable across ages once you filter to
players with at least 500 minutes. This likely happens because the 500-minute
cutoff already removes players who are not getting consistent playing time,
which screens out very young developmental players and older fringe players.

These findings are exploratory. The data is from one season only, the top 100
filter is not random, and things like team system and injuries also affect
per-minute stats. Follow-up work could test these patterns more formally.

# Reflection

Working on this project showed us why normalizing stats matters. Raw totals can
be misleading --- a player with more minutes will almost always have higher
counting numbers, so dividing by minutes levels the comparison.

Combining three separate files into one report also made us think about
consistency. The original files used different variable names and formatting, so
part of the work was standardizing everything so comparisons made sense. Using
shared helper functions (`make_panel` and `make_position_figure`) helped with
that --- instead of copying the same plot code multiple times, we wrote it once
and reused it.

Keeping all the code in the appendix and the body focused on findings also made
the final report much cleaner to read.

\newpage

# Author Contributions {.unnumbered}

**Rishvanth Amsaraj** --- Forwards analysis (data filtering, normalization, and
figure), combining all three position files into the final QMD, patchwork
figure layout, shared helper functions, narrative text, and PCIP documentation.

**Vishnu Bombesetti** --- Centers analysis (data filtering, normalization, and
figure), and the repository README.md file.

**Samarth Puri** --- Guards analysis (data filtering, normalization, and
figure), and the repository PLAN.md file.

The overall analysis approach, filtering criteria, normalization method, and
report structure were decided together as a team.

\newpage

# Appendix 1: PCIP Written Plan {.unnumbered}

The PCIP System (Plan, Code, Implement, Present) was applied to both the data
analysis and the coding work in this project.

## Phase 1 --- Plan {.unnumbered}

**Research question:** How does player age relate to six per-minute performance
statistics across three NBA position groups (Guards, Forwards, Centers)?

**Data source:** `nba-player-stats-2021.csv` --- 2020--21 NBA season, one row
per player.

**Issues identified:**

- Stats are season totals, so players with more minutes look better without
  normalization. Dividing by `mp` fixes this.
- Some players have very few minutes, making their per-minute rates unreliable.
  A 500-minute filter removes these cases.
- The three original files used different variable names and inconsistent
  formatting, which needed to be unified.

**Steps planned:**

1. Load `nba-player-stats-2021.csv` once in the setup chunk.
2. Write shared `make_panel()` and `make_position_figure()` helpers to avoid
   repeating plot code for each position group.
3. For each group: filter by position and `mp > 500`, select the needed
   columns, divide stats by `mp`, and cap at 100 rows.
4. Produce a 2x3 patchwork figure per group with a shared legend.
5. Write narrative for each section and a cross-position discussion.
6. Set up YAML to match the assignment template (TOC, numbered sections,
   1-inch margins, 11pt font, `echo: false` globally).
7. Move all code to a Code Appendix.

## Phase 2 --- Code {.unnumbered}

All code follows the **Tidyverse Style Guide** (https://style.tidyverse.org/).
Key practices:

- Snake\_case for all variable names.
- Pipe operator (`|>`) for data transformation steps.
- Spaces around operators and after commas.
- One `ggplot2` verb per line with consistent indentation.
- Each chunk starts with a comment naming the primary author and reviewer.

## Phase 3 --- Implement {.unnumbered}

The code was run in R through Quarto. We checked that filtering produced the
right rows for each position group, that all six figures rendered correctly,
and that cross-references resolved properly in the PDF.

## Phase 4 --- Present {.unnumbered}

The final PDF was produced by rendering `NBA_Age_vs_Performance.qmd`. The
report includes a table of contents, three cross-referenced figures with
captions and alt text, narrative in every section, a cross-position discussion,
a reflection, author contributions, this PCIP plan, and a code appendix.

\newpage

# Appendix 2: Code Appendix {.unnumbered}

```{r code-appendix, echo=TRUE, eval=FALSE}
# ==========================================================
# Style Guide: Tidyverse Style Guide
# https://style.tidyverse.org/
# ==========================================================

# ── Libraries ──────────────────────────────────────────────
library(tidyverse)
library(ggplot2)
library(patchwork)
library(knitr)
library(kableExtra)

# ── Helper: single scatter panel ──────────────────────────
# Primary Authors: Rishvanth Amsaraj, Vishnu Bombesetti, Samarth Puri
# Reviewer: All team members
make_panel <- function(df, y_var, y_label, panel_title) {
  ggplot(df, aes(x = age, y = .data[[y_var]], color = pos)) +
    geom_point(alpha = 0.75, size = 1.6) +
    labs(
      title  = panel_title,
      x      = "Age",
      y      = y_label,
      color  = "Position"
    ) +
    theme_minimal(base_size = 9) +
    theme(
      plot.title      = element_text(hjust = 0.5, size = 9, face = "bold"),
      legend.position = "bottom",
      legend.key.size = unit(0.35, "cm"),
      legend.text     = element_text(size = 7),
      legend.title    = element_text(size = 7)
    )
}

# ── Helper: 2x3 patchwork figure for a position group ─────
# Primary Authors: Rishvanth Amsaraj, Vishnu Bombesetti, Samarth Puri
# Reviewer: All team members
make_position_figure <- function(df) {
  p1 <- make_panel(df, "pts",       "Points / Min",        "Points")
  p2 <- make_panel(df, "ast",       "Assists / Min",       "Assists")
  p3 <- make_panel(df, "orb",       "Off. Rebounds / Min", "Off. Rebounds")
  p4 <- make_panel(df, "drb",       "Def. Rebounds / Min", "Def. Rebounds")
  p5 <- make_panel(df, "fgpercent", "FG% / Min",           "Field Goal %")
  p6 <- make_panel(df, "tov",       "Turnovers / Min",     "Turnovers")

  ((p1 | p2) / (p3 | p4) / (p5 | p6)) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
}

# ==========================================================
# Section 3: Guards (PG and SG)
# Primary Author: Samarth Puri | Reviewer: Rishvanth Amsaraj
# ==========================================================
playerstats <- read.csv("nba-player-stats-2021.csv")

# Filter to guards with > 500 mp; select analysis columns
guard_data <- playerstats |>
  filter(pos == "PG" | pos == "SG") |>
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, ast, orb, drb, fgpercent, tov)

# Normalize each statistic by minutes played
cols_to_normalize <- c("pts", "ast", "orb", "drb", "fgpercent", "tov")
for (colname in cols_to_normalize) {
  guard_data[[colname]] <- round(
    guard_data[[colname]] / guard_data[["mp"]], 6
  )
}
guard_data <- head(guard_data, 100)

make_position_figure(guard_data)

# ==========================================================
# Section 4: Forwards (PF and SF)
# Primary Author: Rishvanth Amsaraj | Reviewer: Vishnu Bombesetti
# ==========================================================

# Filter to forwards with > 500 mp; select analysis columns
forward_data <- playerstats |>
  filter(pos == "PF" | pos == "SF") |>
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, ast, orb, drb, fgpercent, tov)

# Normalize each statistic by minutes played
for (colname in cols_to_normalize) {
  forward_data[[colname]] <- round(
    forward_data[[colname]] / forward_data[["mp"]], 6
  )
}
forward_data <- head(forward_data, 100)

make_position_figure(forward_data)

# ==========================================================
# Section 5: Centers (C)
# Primary Author: Vishnu Bombesetti | Reviewer: Rishvanth Amsaraj
# ==========================================================

# Filter to centers with > 500 mp; select analysis columns
center_data <- playerstats |>
  filter(pos == "C") |>
  filter(mp > 500) |>
  select(player, pos, age, mp, pts, ast, orb, drb, fgpercent, tov)

# Normalize each statistic by minutes played
for (colname in cols_to_normalize) {
  center_data[[colname]] <- round(
    center_data[[colname]] / center_data[["mp"]], 6
  )
}
center_data <- head(center_data, 100)

make_position_figure(center_data)
```

## References {.unnumbered}

Basketball Reference. (2021). *NBA player stats --- 2020--21 season* [Dataset].
Retrieved from https://www.basketball-reference.com

Tukey, J. W. (1977). *Exploratory data analysis*. Addison-Wesley.

Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D., François, R.,
Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T. L.,
Miller, E., Bache, S. M., Müller, K., Ooms, J., Robinson, D., Seidel, D. P.,
Spinu, V., Takahashi, K., Vaughan, D., Wilke, C., Woo, K., & Yutani, H. (2019).
Welcome to the tidyverse. *Journal of Open Source Software*, *4*(43), 1686.
https://doi.org/10.21105/joss.01686
